---
title: "Statistical Comparison"
output: html_document
date: "2023-12-22"
---

```{r}
library (readr)
library (dplyr)
library(ggplot2)
library(tidyr)
data <- read.csv("~/Module6/BPV Indices Results edit 1.csv")
```

## Add column to the imported csv to add information of injury category- paraplegic and teraplegic

```{r cars}
#Adding a column with values paraplegic and tetraplegic category
data <- data %>% mutate(Injury.Category = case_when(
      substr(Participant.characteristics.NLI, 1, 1) == "C" ~ "tetraplegic",
      TRUE ~ "paraplegic"))

```

## Selecting columns with mean data for SBP, DBP and Pulse and Plotting

```{r}

#Reshape data for plotting Mean SBP
data_mean_sbp <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Mean.SBP"))
ggplot(data_mean_sbp, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "Mean SBP", title = "Comparison of Mean SBP by Injury Category") +
  theme_minimal()


#Reshape data for plotting Mean DBP
data_mean_dbp <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Mean.DBP"))
ggplot(data_mean_dbp, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "Mean DBP", title = "Comparison of Mean DBP by Injury Category") +
  theme_minimal()


#Reshape data for plotting Mean Pulse
data_mean_pulse <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Mean.PULSE"))
ggplot(data_mean_pulse, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "Mean Pulse", title = "Comparison of Mean Pulse by Injury Category") +
  theme_minimal()


```
## Selecting columns with SD data for SBP, DBP and Pulse and Plotting

```{r}
#Reshape data for plotting SD SBP
data_sd_sbp <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Standard.Deviation.SBP"))
ggplot(data_sd_sbp, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "SD SBP", title = "Comparison of SD SBP by Injury Category") +
  theme_minimal()

#Reshape data for plotting SD DBP
data_sd_dbp <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Standard.Deviation.DBP"))
ggplot(data_sd_dbp, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "SD DBP", title = "Comparison of SD DBP by Injury Category") +
  theme_minimal()

#Reshape data for plotting SD Pulse
data_sd_pulse <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Standard.Deviation.PULSE"))
ggplot(data_sd_pulse, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "SD Pulse", title = "Comparison of SD pulse by Injury Category") +
  theme_minimal()
```


## Selecting columns with CoV data for SBP, DBP and Pulse and Plotting
```{r}

#Reshape data for plotting CoV SBP
data_cov_sbp <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Coefficient.of.Variation.SBP"))
ggplot(data_cov_sbp, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "CoV SBP", title = "Comparison of CoV SBP by Injury Category") +
  theme_minimal()

#Reshape data for plotting CoV DBP
data_cov_dbp <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Coefficient.of.Variation.DBP"))
ggplot(data_cov_dbp, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "CoV DBP", title = "Comparison of CoV DBP by Injury Category") +
  theme_minimal()

#Reshape data for plotting CoV Pulse
data_cov_pulse <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Coefficient.of.Variation.PULSE"))
ggplot(data_cov_pulse, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "CoV Pulse", title = "Comparison of CoV Pulse by Injury Category") +
  theme_minimal()
```
## Selecting columns with ARV data for SBP, DBP and Pulse and Plotting
```{r}

#Reshape data for plotting ARV SBP
data_arv_sbp <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Average.Real.Variability.SBP"))
ggplot(data_arv_sbp, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "ARV SBP", title = "Comparison of ARV SBP by Injury Category") +
  theme_minimal()

#Reshape data for plotting ARV DBP
data_arv_dbp <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Average.Real.Variability.DBP"))
ggplot(data_arv_dbp, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "ARV DBP", title = "Comparison of ARV DBP by Injury Category") +
  theme_minimal()

#Reshape data for plotting ARV Pulse
data_arv_pulse <- tidyr::gather(data, key = "Variable", value = "Value", starts_with("Average.Real.Variability.PULSE"))
ggplot(data_arv_pulse, aes(x = Injury.Category, y = Value, fill = Variable)) +
  geom_boxplot() +
  labs(x = "Injury Category", y = "ARV Pulse", title = "Comparison of ARV Pulse by Injury Category") +
  theme_minimal()

```
